---
title: "Reproducible Research - Project 1"
author: "Hayden Roche"
date: "December 7, 2017"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
setwd("/Users/hroche/Coursera/Coursera/")
Project_data <- read.csv("activity.csv")
knitr::opts_chunk$set(echo = TRUE)
```


##Total Number of Steps Taken Each Day
```{r dailysteps}
Histo <-  Project_data %>%
  group_by(date) %>%
  summarise(Total_Steps = sum(steps))

g <- ggplot(Histo, aes(x = Total_Steps)) +
  geom_histogram(fill = "orange") +
  labs(title = "Steps Per Day", x = "Date", y = "Frequency")
print(g)
```


##Mean and median number of steps taken each day
```{r mean_and_median}
Mean <- mean(Histo$Total_Steps, na.rm = TRUE)
Mean  
Median <- median(Histo$Total_Steps, na.rm = TRUE)
Median
```

##Time series: average number of steps per day
```{r time_series}
Time_Series <- Project_data %>%
  group_by(interval) %>%
  filter(!is.na(steps)) %>%
  summarise(Average = mean(steps, na.rm = TRUE))

g <- ggplot(Time_Series, aes(x = interval, y = Average)) +
  geom_line(color="purple") +
  labs(title = "Average Number of Steps Taken by Interval", y = "Average Daily Steps")
print(g)
```


This 5-minute interval on average contains the maximum number of steps
```{r interval_max}
Time_Series[which.max(Time_Series$Average),]
```


##Imputing Values
```{r impute}
sum(is.na(Project_data$steps))

Imputing_Values <- Time_Series$Average[match(Project_data$interval, Time_Series$interval)]
Fill_Tha_Data <- transform(Project_data, steps = ifelse(is.na(Project_data$steps), yes = Imputing_Values, no = Project_data$steps))

Imputed_Histo <-  Fill_Tha_Data %>%
  group_by(date) %>%
  summarise(Total_Steps = sum(steps))
g <- ggplot(Imputed_Histo, aes(x = Total_Steps)) +
  geom_histogram(fill = "grey") +
  labs(title = "Steps Per Day", x = "Date", y = "Frequency")
print(g)
```

This is where I calculated the difference between the imputed mean/median and the original estimation of mean/median
```{r imputed_mean_median}
Imputed_Mean <- mean(Imputed_Histo$Total_Steps)
Imputed_Mean
Imputed_Median <- median(Imputed_Histo$Total_Steps)
Imputed_Median

Mean_Difference <- Imputed_Mean - Mean
Mean_Difference
Median_Difference <- Imputed_Median - Median
Median_Difference
Total_Step_Difference <- sum(Imputed_Histo$Total_Steps) - sum(Histo$Total_Steps)
Total_Step_Difference
```

##Panel Plot: Average Number of Steps Taken (Weekend vs. Weekdays)
```{r weekendvweekday}
Project_data$date <- as.Date(strptime(Project_data$date, format = "%Y-%m-%d"))
Project_data$segment <- sapply(Project_data$date, function(x) {
  if(weekdays(x) == "Saturday" | weekdays(x) == "Sunday")
  {y <- "Weekend"} else
  {y <- "Weekday"}
  y
})
 
By_Date_Steps <- aggregate(steps~interval + segment, Project_data, mean, na.rm = TRUE)
g <- ggplot(By_Date_Steps, aes(x = interval, y = steps)) +
  geom_line() +
  labs(title = "Steps Per Day", x = "Date", y = "Frequency") +
  facet_wrap(~segment, ncol = 1, nrow = 2)
print(g)
```

